---
title: "using qckitfastq package"
author: "Wenyue Xing"
package: qckitfastq
vignette: >
  %\VignetteIndexEntry{quality control analysis and visualization using qckitfastq }
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



# I. Pre-processing of FASTQ files

We first install the package from github by using `devtools`

```{R,include=FALSE}

library(devtools)

install_github("compbiocore/qckitfastq",build_vignettes=TRUE)


```

Then we specify the path to the FASTQ file and run analysis and visualization functions on it. 
Here, for example, we use a system file with 25,000 reads as an example. 

```{r}
library(qckit)
infile <- system.file("extdata", "10^5_reads_test.fq.gz", package = "qckit")

```
To open FASTQ files, we can replace the infile argument with path to the FASTQ file. 

```{r}

#infile <- "Data/10^5_reads_test.fq.gz"

```

For some functions based on `seqTools`, we will first process the file through seqTools and then let qckitfastq functions proceed with the analysis. 

```{r}

fseq <- seqTools::fastqq(infile)

```

# II. Metrics of FASTQ quality control


## 1. data dimension

*dimensions* function obtains the number of reads and number of positions from a `seqTools` processed FASTQ file. The sample file has 25,000 reads and 100 positions. This section allows us to examine whether the sequence length are uniform throughout the entire file. The results from this section can also help with further analysis like *overrepresented_kmer*, where dimensionality of data is important for calculation. 

```{r}


nc <- dimensions(fseq,"positions")
nc

nr <- dimensions(fseq,"reads")
nr
```

*plot_sequence_length* function generates distribution plot of the length of all reads. The generated plot would show the sequence length of all the sequences throughout the file. The plot is considered an indication of good data quality is all sequences have the same sequence length with no deviations. The following plot shows that all reads in the file have sequence length of 100. 

```{r}
plot_sequence_length(fseq,writefile=FALSE)


```

## 2.  per base quality score statistics

*basic_statistics* function calculates the quality score per sequence statistics including mean, median and quantiles. We will use the result to create a quality score distribution per position plot to present the overall dispersion and concentration of quality score per position of the FASTQ file.   

```{r}

bs <- basic_stat(infile,FALSE)
knitr::kable(head(bs))
plot_quality_score(bs,writefile=FALSE)
```
*plot_quality_score* function generates boxplot of quality score distribution per base.In this section, we would like to examine whether certain portion of the sequences have generally low quality score. If a subset of seqences have quality score generally small(<30) 

## 3. nucleotide sequence content

*sequence_content* calculates the total number of each nucleotide sequence content per sequence throughout the file. 
```{r}

scA <- sequence_content(fseq, content = "A",writefile=FALSE)
scA
```

*plot_sequence_conent* use the *sequence_content* function and plot the percentage of all nucleotide sequence content per position. The plot would be considered an indication of good data quality when the percentage of each nucleotide sequence content is not significantly larger than other nucleotide sequence contents. In this section, we examine whether each nucleotide sequence content are distributed evenly at all bases. The data is problematic if any sequence bias greatly from other sequence content in any subset of sequences. 

```{r}

plot_sequence_content(fseq,nr,nc)

```


## 4. GC content


Function *GC_content* utilizes the result from the C++_calling function *GC_per_read* and produce the data frame that is more suitable for the plotting purpose as well as for saving to data file for later access. The result is the GC content percentage per read and will be saved in resulting table. 

```{r}

gc_df <- GC_content(infile)

knitr::kable(head(gc_df))


```


The *plot_GC_content_plot* function takes the input from previous function and generates plot of distribution of GC nucleotide sequence content percentage. The graph would be an indication of good data quality when the  GC percentage concentrated around 30 to 50 and the results do not differ among different bases. 


```{r}

plot_GC_content(nc,gc_df)


```


## 5. Per read sequence quality score

*plot_perseq_quality* function extracts the mean quality score per read and generate a histogram of this statistics.It takes the input of the path to the gzipped FASTQ files. It also enables plot saving if set *writefile* to TRUE and input a *prefix* value to customize name of saved file. The histogram would be considered an indication of good data quality when the majority of reads have high quality score(around 30). If a significant portion of read are with low quality score, say less than 20, then data is problematic. 


```{r}

plot_perseq_quality(infile,writefile=FALSE)

```

## 6. Kmers count per base

*kmer_count* function produces the per position kmer count with given path to the FASTQ file and the kmer length specified. 

```{r}

km <- Kmer_count(infile,k=6)
knitr::kable(km[1:20,1:10])

```

## 7. Overrepresented Kmers
The function generates list of overrepresented kmers with its maximum obs/exp ratio and the position of maximum obs/exp ratio. 


```{r}

overkm <-qckit::overrep_kmer(infile,7,nc,nr)
knitr::kable(head(overkm,n=10))


```

## 8. Overrepresented Sequence

Taking in the output generated from *cal_over_rep_seq*, function *overrepresented_sequence* produce the overrepresented sequence table by selecting the unique sequences that have counts larger than 0.1% of the total reads of the data file. The results would be displayed as table with decreasing order of counts.


```{r}

overrep_seq <- qckit::overrep_sequence(infile,nr,FALSE,"test")
knitr::kable(head(overrep_seq,n = 5))

```

Function *overrepresented_plot* would take the overrepresented sequence table as input and present a density plot of the counts and mark the top 5 overrepresented sequence in red. 


```{r}


plot_overrep_seq(overrep_seq,FALSE)


```
